---
title: "Adverse effect of bexaroten therapy in T-lymphom "
author: "Jiri Baloun"
date: "Compiled on: **2025-04-25**  <br> Last updated **`r format(Sys.time(), '%Y-%m-%d')`**"
output:
  html_document:
    code_folding: hide
    df_print: paged
    theme: yeti
    highlight: tango
    toc: yes
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: true
  pdf_document:
    fig_caption: yes
    toc: yes
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style type="text/css">
.title {
  text-align: center;
}
.author {
  font-size: 30px;
  text-align: center;
  font-style: italic;
  font-weight: 300;
}
.date {
  font-weight: 300;
}
h1, h2, h3, h4 {
    font-weight: 600;
    margin-top: 2em;
}
body {text-align: justify}
blockquote {
    padding: 10px 20px;
    margin: 0 0 20px;
    font-size: 14px;
    border-left: 5px solid #eee;
}
p {
  text-indent: 2rem;
}
table {
    width:100%;
}
.main-container {
  max-width: 100% !important;
  margin: auto;
}
img{
    height: 100%;
    width: 100%;
    object-fit: cover;
}
.boxBorder {
     border: 2px solid #990066;
     padding: 10px;
     outline: #990066 solid 5px;
     outline-offset: 5px;
   }
</style>
```
```{r setup, fig.align = 'center', include=FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    message = FALSE,
    warning = FALSE,
    knitr.kable.NA = ''  # if it does not work, use this  options(knitr.kable.NA = '')
)
options(knitr.kable.NA = '')

# Functions
## loading  
pacman::p_load(update = FALSE,
  tidyverse,readxl, data.table, magrittr, Rmisc, future, 
               purrr, easystats, rio, janitor, ggthemes, car, labelled,testthat,
               lubridate, dlookr, gtsummary, skimr, summarytools,kableExtra,
               psych, smplot2, car, broom, sjPlot, glue, GGally,
               parallelly,tidymodels, corrr, factoextra, FactoMineR) 
## function specification ----
conflicted::conflicts_prefer(
  janitor::remove_empty,
  dplyr::filter,
  dplyr::mutate,
  dplyr::rename,
  dplyr::summarize,
  dplyr::summarise,
  dplyr::select,
  purrr::map,
  tidyr::extract,
  stats::chisq.test,
  base::intersect,
  base::setdiff,
  dplyr::last,
  dplyr::first, 
  dplyr::between,
  corrr::correlate,
  purrr::set_names,
  dplyr::desc
)

## Manual 
# skimr
my_skim <- skimr::skim_with(numeric = sfl(median = ~ median(., na.rm = TRUE),
                                          mad = ~ mad(., na.rm = TRUE)), 
                            append = TRUE)

# Load objects
load("markD_02.RData")

# Load cores
no_cores <- parallelly::availableCores() -1

# Load functions
list.files("F:/Analysis/github/project-ettler01/scripts/functions/", pattern="*.*", full.names=TRUE) %>% str_subset( "OBJ", negate = TRUE) %>%  map(~source(.))

# Parameter selection ----
param_sel_2 <- c("initials", "age", "sex", "bmi", "ps_ecog", "stage_early",
                 "first_syst_th", "response_achieved", "thyroid_disease_before",
                 "dyslipidemia_before", "monotherapy", "ttnt", "ttnt_achieved",
                 "response_time_to", "response_duration",	"progression",
                 "treatment_duration", "discontinuation_reason", "ae_grade_3_4"
)

variab_dep_01 <- c("discontinued_due_to_ae", "ae_complete_grade_3_4", 
                    "ae_hyperTAG_grade_3_4", "ae_liver",	"ae_hemato")
variab_indep_01 <- c("age", "sex", "bmi", "ps_ecog", "stage_early", "first_syst_th",
                     "response_achieved", "dyslipidemia_before", 
                     "thyroid_disease_before", "monotherapy"  
)

variab_dep_02 <- c("stage_early", "response_achieved", 
                   "ttnt", "ttnt_achieved",
                   "response_time_to",                 # only non-zero patients
                   "response_duration", "progression",
                   "treatment_duration", "discontinuation_reason"
)
variab_indep_02 <- c("age", "sex", "bmi", "ps_ecog", "first_syst_th", "ae_grade_3_4",
                     "monotherapy")



# Object processing  
# d02 <- d01 |> select(any_of(param_sel))
# d02_binary <- d02 |>
#   select(-initials, -age, -bmi, -ps_ecog) |>
#   mutate(sex = case_when(
#     sex == "F" ~ 0,
#     TRUE ~ 1
#   )) 


```


# Questions  
- [ ] In column $response_achieved$ there is value $x$. What does it mean?  
- [ ] The same is in the column $discontinued_due_to_ae$. What does the $x$ mean?  
- [ ] I removed all $lost of FU$.  
- [ ] Should I use $ae_hyperTAG_any$ (column AP) instead of $ae_complete_grade_3_4$ (column AN) in question I.A.2.? In the $ae_hyperTAG_any$ (column AP), there is the selection of any *hyperTAG* in adverse effect.  


# Notes  
* $ae_hyperTAG_any$ is perfectly predicted by $thyroid_disease_before"$. In other words, for that pair of variables, the logistic regression model fits too perfectly — the predicted probabilities are 0 or 1 without uncertainty.













# Session info

<details>

<summary>Platform</summary>

```{r session_info_1}
df_session_platform <- devtools::session_info()$platform %>% 
  unlist(.) %>% 
  as.data.frame(.) %>% 
  rownames_to_column(.)

colnames(df_session_platform) <- c("Setting", "Value")

kable(
  df_session_platform, 
  booktabs = T, 
  align = "l",
  caption = "(ref:Reproducibility-SessionInfo-R-environment-title)", # complete caption for main document
  caption.short = " " # "(ref:Reproducibility-SessionInfo-R-environment-caption)" # short caption for LoT
) %>% 
  kable_styling(full_width = F,
                latex_options = c(
                  "hold_position" # stop table floating
                ),
                bootstrap_options = c("striped", "hover", "condensed", "responsive")
  ) 

```

</details>

<details>

<summary>Used packages</summary>

```{r session_info_2}
df_session_packages <- devtools::session_info()$packages %>% 
  as.data.frame(.) %>% 
  filter(attached == TRUE) %>% 
  dplyr::select(loadedversion, date) %>% 
  rownames_to_column

colnames(df_session_packages) <- c("Package", "Loaded version", "Date")

rows <- seq_len(nrow(df_session_packages) %/% 2)
l_session_packages <- list(df_session_packages[rows,1:3],  
                           matrix(numeric(), nrow=0, ncol=1),
                           df_session_packages[-rows, 1:3])

kable(
  l_session_packages, 
  row.names = T,
  booktabs = T, 
  align = "l",
  caption = "(ref:Reproducibility-SessionInfo-R-packages-title)", # complete caption for main document
  caption.short = " " # "(ref:Reproducibility-SessionInfo-R-packages-caption)" # short caption for LoT
) %>% kable_styling(full_width = F,
                    latex_options = c(
                      "hold_position" # stop table floating
                    ),bootstrap_options = c("striped", "hover", "condensed", "responsive") 
) 
```

</details>

<br> <br> <br> <br> <br> <br>

::: {.tocify-extend-page data-unique="tocify-extend-page" style="height: 0;"}
:::

