---
title: "Adverse effect of bexaroten therapy in T-lymphom "
author: "Jiri Baloun"
date: "Compiled on: **2025-04-01**  <br> Last updated **`r format(Sys.time(), '%Y-%m-%d')`**"
output:
  html_document:
    code_folding: hide
    df_print: paged
    theme: yeti
    highlight: tango
    toc: yes
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: true
  pdf_document:
    fig_caption: yes
    toc: yes
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style type="text/css">
.title {
  text-align: center;
}
.author {
  font-size: 30px;
  text-align: center;
  font-style: italic;
  font-weight: 300;
}
.date {
  font-weight: 300;
}
h1, h2, h3, h4 {
    font-weight: 600;
    margin-top: 2em;
}
body {text-align: justify}
blockquote {
    padding: 10px 20px;
    margin: 0 0 20px;
    font-size: 14px;
    border-left: 5px solid #eee;
}
p {
  text-indent: 2rem;
}
table {
    width:100%;
}
.main-container {
  max-width: 100% !important;
  margin: auto;
}
img{
    height: 100%;
    width: 100%;
    object-fit: cover;
}
.boxBorder {
     border: 2px solid #990066;
     padding: 10px;
     outline: #990066 solid 5px;
     outline-offset: 5px;
   }
</style>
```
```{r setup, fig.align = 'center', include=FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    message = FALSE,
    warning = FALSE,
    knitr.kable.NA = ''  # if it does not work, use this  options(knitr.kable.NA = '')
)
options(knitr.kable.NA = '')

# Functions
## loading  
pacman::p_load(update = FALSE,
  tidyverse,readxl, data.table, magrittr, Rmisc, future, 
               purrr, easystats, rio, janitor, ggthemes, car, labelled,testthat,
               lubridate, dlookr, gtsummary, skimr, summarytools,kableExtra,
               psych, smplot2, car, broom, sjPlot, glue, GGally,
               parallelly,tidymodels, corrr, factoextra, FactoMineR, ggpubr) 
## function specification ----
conflicted::conflicts_prefer(
  janitor::remove_empty,
  dplyr::filter,
  dplyr::mutate,
  dplyr::rename,
  dplyr::summarize,
  dplyr::summarise,
  dplyr::select,
  purrr::map,
  tidyr::extract,
  stats::chisq.test,
  base::intersect,
  base::setdiff,
  dplyr::last,
  dplyr::first, 
  dplyr::between,
  corrr::correlate,
  purrr::set_names,
  dplyr::desc
)

## Manual 
# skimr
my_skim <- skimr::skim_with(numeric = sfl(median = ~ median(., na.rm = TRUE),
                                          mad = ~ mad(., na.rm = TRUE)), 
                            append = TRUE)

# Load objects
load("markD_01.RData")

# Load cores
no_cores <- parallelly::availableCores() -1

# Load functions
list.files("F:/Analysis/github/project-ettler01/scripts/functions/", pattern="*.*", full.names=TRUE) %>% str_subset( "OBJ", negate = TRUE) %>%  map(~source(.))

# Parameter selection ----
param_sel <- c("initials", "age", "sex", "bmi", "ae_grade_3_4",
               "ae_hyperTAG_grade_3_4","discontinued_due_to_ae", "ae_liver",
               "ae_hemato", "ps_ecog", "stage_early", "first_syst_th",
               "thyroid_disease_before", "monotherapy", "response_achieved",
               "dyslipidemia_before")


# Object processing  
d02 <- d01 |> select(any_of(param_sel))
d02_binary <- d02 |>
  select(-initials, -age, -bmi, -ps_ecog) |>
  mutate(sex = case_when(
    sex == "F" ~ 0,
    TRUE ~ 1
  )) 


```

# Questions

-   [ ] In "AE Grade 3 / 4" (AN column of the provided excel), there is complete information of AE with hyperTAG. However, there is a column with the same name and binominal values. Could I use this column?
-   [ ] Is the occurance of hyperTAG before bexarten the same as dyslipidimia occurence? Column AK in the excel table
-   [ ] Is the EA connected to monotherapy? If monotherapy works, AE also occures?

# Exploratory Data Analyses (EDA) {.tabset .tabset-fade .tabset-pills}

## Overview {.unnumbered}

```{r eda_skimr, results="asis"}
d02 |> 
  select(-initials) |>  
  mutate(across(
    .cols = -c(bmi, age, ps_ecog), 
    .fns = as.factor
  )) |> 
  my_skim()
```


## Proportional Bar plot {.unnumbered}

```{r eda_01}
d02 |> 
  select(-initials, -age, -bmi) |> 
  mutate(across(everything(),as.factor)) |> 
  pivot_longer(cols = everything()) |> 
  na.omit() |> 
  ggplot(aes(x = name, fill = as.factor(value))) +
  geom_bar(position = "fill") +  # Proportions
  labs(y = "Proportion", fill = "Outcome") +
  theme_sjplot2()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  paletteer::scale_fill_paletteer_d("waRhol::marilyn_orange_62") +
  NULL
```

## Proportional Baloon plot {.unnumbered}

```{r eda_02}
d02_binary |> 
  pivot_longer(cols = c("ae_grade_3_4","ae_hyperTAG_grade_3_4","ae_liver",
                        "discontinued_due_to_ae", "ae_liver", "ae_hemato"),
               names_to = "ae_name",
               values_to = "ae_value") |> 
  na.omit() |> 
  group_by(ae_name, ae_value) |> 
  summarise(across(sex:response_achieved, sum), .groups = "drop") |>
  mutate(
    ae_name = as.factor(ae_name),
    ae_value = case_when(
      ae_value == 0 ~ "no",
      ae_value == 1 ~ "yes"
    ),
    ea = paste(ae_name, ae_value, sep = "_")
  ) |>
  select(-ae_name, -ae_value) |> 
  column_to_rownames("ea") |> 
  ggballoonplot(fill = "value") +
  scale_fill_viridis_c(option = "C")
```

## Correlation {.tabset .unnumbered}
### Tetrachoric analyses {.unnumbered}
```{r eda_03_tetrachoric}
res_tetra <- psych::tetrachoric(
  d02_binary
  )


corrplot::corrplot(res_tetra$rho, method = 'square', tl.col = 'black',
                   order = 'AOE', type = 'lower', addCoef.col = 'black',
                   col=colorRampPalette(c("blue","white","red"))(200),
                   diag = FALSE)
```
  

### Pearson correlation analyses {.unnumbered}
```{r eda_04_pearson}
d02_binary |> 
  correlate()  |> 
  shave() |> 
  rplot(print_cor=TRUE,
        colours = colorRampPalette(c("blue","white","red"))(200)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```
  
### Clustering {.tabset .unnumbered}  

#### Heatmap {.unnumbered}

```{r eda_05_heatmap}
fig_heatmap_bin_01 <- d02_binary |> 
  pivot_longer(cols = c("ae_grade_3_4","ae_hyperTAG_grade_3_4","ae_liver",
                        "discontinued_due_to_ae", "ae_liver", "ae_hemato"),
               names_to = "ae_name",
               values_to = "ae_value") |> 
  na.omit() |> 
  relocate(ae_value) |> 
  group_by(ae_name) |> 
  nest() |> 
  mutate(fig = map2(data, ae_name, ~heatmap_bin(df = .x, name_ae = .y)))

walk(fig_heatmap_bin_01$fig, print)
```



#### Multiple visualization {.unnumbered}

```{r eda_06_alter_venn}
ComplexUpset::upset(d02_binary, intersect = d02_binary |> names())
```



# Statistics
## Multiple correspondence analyses

```{r stat_01_mca}
res_mca <- d02 |> select(-initials, -age, -bmi) |> na.omit() |> 
  mutate(across(everything(),as.factor)) |> 
  MCA(graph = FALSE)
fviz_mca_var(res_mca, repel = TRUE, ggtheme = theme_minimal(),
             col.var = "contrib",
             # choice = "mca.cor",
             gradient.cols = c("#bdbdbd", "#ffeda0", "#FC4E07"))
fviz_screeplot(res_mca)
fviz_ellipses(res_mca, c("sex", "ae_grade_3_4"),
              geom = "point")
dimdesc(res_mca, axes = c(1,2))

```



## Univariate & Multivariable logistic regression

```{r stat_uva_mva}
cluster <- new_cluster(no_cores)
cluster_library(cluster, "MASS") |> cluster_library("data.table") |> 
  cluster_library("tidyverse") |> cluster_library("gtsummary")
cluster_copy(cluster, 'tab_logist_anal') |> 
  cluster_copy('tab_logist_anal_uva')

res_logist_tab_01 <- d02 |>
  mutate(sex = case_when(
    sex == "F" ~ 0,
    TRUE ~ 1
  )) |> 
  pivot_longer(cols = c("ae_grade_3_4","ae_hyperTAG_grade_3_4","ae_liver",
                        "discontinued_due_to_ae", "ae_liver", "ae_hemato"),
               names_to = "ae_name",
               values_to = "ae_value") |> 
  na.omit() |> 
  relocate(ae_value) |> 
  group_by(ae_name) |> 
  nest() |> 
  multidplyr::partition(cluster = cluster) |> 
  mutate(mod_01 = map2(data, ae_name, 
                    ~tab_logist_anal(data = .x, 
                                     dependent_variable = .y,
                                     independent_variable = c("bmi", "sex", "age"))),
         mod_02 = map2(data, ae_name, 
                       ~tab_logist_anal(data = .x, 
                                        dependent_variable = .y,
                                        independent_variable = c("bmi", "sex"))),
         mod_03 = map2(data, ae_name, 
                       ~tab_logist_anal(data = .x, 
                                        dependent_variable = .y,
                                        independent_variable = c("bmi", "age"))),
         mod_04 = map2(data, ae_name, 
                       ~tab_logist_anal_uva(data = .x, 
                                        dependent_variable = .y,
                                        independent_variable = c("stage_early","ps_ecog",
                                                                 "first_syst_th", "response_achieved",
                                                                 "monotherapy", "dyslipidemia_before"))),
         mod_05 = map2(data, ae_name, 
                       ~tab_logist_anal(data = .x, 
                                        dependent_variable = .y,
                                        independent_variable = c("bmi", "sex", "age",
                                                                 "stage_early"))),
         mod_06 = map2(data, ae_name, 
                       ~tab_logist_anal(data = .x, 
                                        dependent_variable = .y,
                                        independent_variable = c("bmi", "sex", "age",
                                                                 "ps_ecog"))),
         mod_07 = map2(data, ae_name, 
                       ~tab_logist_anal(data = .x, 
                                        dependent_variable = .y,
                                        independent_variable = c("bmi", "sex", "age",
                                                                 "first_syst_th"))),
         mod_08 = map2(data, ae_name, 
                       ~tab_logist_anal(data = .x, 
                                        dependent_variable = .y,
                                        independent_variable = c("bmi", "sex", "age",
                                                                 "response_achieved"))),
         mod_09 = map2(data, ae_name, 
                       ~tab_logist_anal(data = .x, 
                                        dependent_variable = .y,
                                        independent_variable = c("bmi", "sex", "age",
                                                                 "monotherapy"))),
         mod_10 = map2(data, ae_name, 
                       ~tab_logist_anal(data = .x, 
                                        dependent_variable = .y,
                                        independent_variable = c("bmi", "sex", "age",
                                                                 "dyslipidemia_before")))
         ) |> 
  collect()
walk(cluster, ~ .x$kill())
```


### ~bmi + sex + age
```{r stat_mod1, results="asis"}
walk(res_logist_tab_01$mod_01, print)
```

### ~bmi + sex
```{r stat_mod2, results="asis"}
walk(res_logist_tab_01$mod_02, print)

```

### ~bmi + age
```{r stat_mod3, results="asis"}
walk(res_logist_tab_01$mod_03, print)

```

### uvivariate analyses
```{r stat_mod4, results="asis"}
walk(res_logist_tab_01$mod_04, print)

```

### ~bmi + sex + age + StageEarly
```{r stat_mod5, results="asis"}
walk(res_logist_tab_01$mod_05, print)

```


### ~bmi + sex + age + ecog
```{r stat_mod6, results="asis"}
walk(res_logist_tab_01$mod_06, print)

```

### ~bmi + sex + age + FirstSyst
```{r stat_mod7, results="asis"}
walk(res_logist_tab_01$mod_07, print)

```

### ~bmi + sex + age + ResponseAchieved
```{r stat_mod8, results="asis"}
walk(res_logist_tab_01$mod_08, print)

```

### ~bmi + sex + age + monotherapy
```{r stat_mod9, results="asis"}
walk(res_logist_tab_01$mod_09, print)

```


### ~bmi + sex + age + dyslipidemia
```{r stat_mod10, results="asis"}
walk(res_logist_tab_01$mod_10, print)

```







# Session info

<details>

<summary>Platform</summary>

```{r session_info_1}
df_session_platform <- devtools::session_info()$platform %>% 
  unlist(.) %>% 
  as.data.frame(.) %>% 
  rownames_to_column(.)

colnames(df_session_platform) <- c("Setting", "Value")

kable(
  df_session_platform, 
  booktabs = T, 
  align = "l",
  caption = "(ref:Reproducibility-SessionInfo-R-environment-title)", # complete caption for main document
  caption.short = " " # "(ref:Reproducibility-SessionInfo-R-environment-caption)" # short caption for LoT
) %>% 
  kable_styling(full_width = F,
                latex_options = c(
                  "hold_position" # stop table floating
                ),
                bootstrap_options = c("striped", "hover", "condensed", "responsive")
  ) 

```

</details>

<details>

<summary>Used packages</summary>

```{r session_info_2, results = 'asis'}
subset(data.frame(sessioninfo::package_info()), attached==TRUE, c(package, loadedversion, date))
```


```{r}
subset(data.frame(sessioninfo::package_info()), attached==TRUE, c(package, loadedversion, date))
```


</details>

<br> <br> <br> <br> <br> <br>

::: {.tocify-extend-page data-unique="tocify-extend-page" style="height: 0;"}
:::
