---
title: "Adverse effect of bexaroten therapy in T-lymphom "
author: "Jiri Baloun"
date: "Compiled on: **2025-04-25**  <br> Last updated **`r format(Sys.time(), '%Y-%m-%d')`**"
output:
  html_document:
    code_folding: hide
    df_print: paged
    theme: yeti
    highlight: tango
    toc: yes
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: true
  pdf_document:
    fig_caption: yes
    toc: yes
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style type="text/css">
.title {
  text-align: center;
}
.author {
  font-size: 30px;
  text-align: center;
  font-style: italic;
  font-weight: 300;
}
.date {
  font-weight: 300;
}
h1, h2, h3, h4 {
    font-weight: 600;
    margin-top: 2em;
}
body {text-align: justify}
blockquote {
    padding: 10px 20px;
    margin: 0 0 20px;
    font-size: 14px;
    border-left: 5px solid #eee;
}
p {
  text-indent: 2rem;
}
table {
    width:100%;
}
.main-container {
  max-width: 100% !important;
  margin: auto;
}
img{
    height: 100%;
    width: 100%;
    object-fit: cover;
}
.boxBorder {
     border: 2px solid #990066;
     padding: 10px;
     outline: #990066 solid 5px;
     outline-offset: 5px;
   }
</style>
```
```{r setup, fig.align = 'center', include=FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    message = FALSE,
    warning = FALSE,
    knitr.kable.NA = ''  # if it does not work, use this  options(knitr.kable.NA = '')
)
options(knitr.kable.NA = '')

# Functions
## loading  
pacman::p_load(update = FALSE,
               tidyverse, purrr, furrr, easystats, rio, janitor, ggthemes, car,
               gtsummary, skimr, sjPlot, flextable, ggpubr, rstatix, tidymodels,
               psych, paletteer, ComplexHeatmap, future, multidplyr, corrr, 
               factoextra, lmerTest, ggforce, lazyWeave, FactoMineR, sjPlot,
               kableExtra, survival, survminer, ggsurvfit, ggfortify, 
               adjustedCurves
  ) 
## function specification ----
conflicted::conflicts_prefer(
  janitor::remove_empty,
  dplyr::filter,
  dplyr::mutate,
  dplyr::rename,
  dplyr::summarize,
  dplyr::summarise,
  dplyr::select,
  purrr::map,
  tidyr::extract,
  stats::chisq.test,
  base::intersect,
  base::setdiff,
  dplyr::last,
  dplyr::first, 
  dplyr::between,
  corrr::correlate,
  purrr::set_names,
  dplyr::desc
)

## Manual 
# skimr
my_skim <- skimr::skim_with(numeric = sfl(median = ~ median(., na.rm = TRUE),
                                          mad = ~ mad(., na.rm = TRUE)), 
                            append = TRUE)

# Load objects
load("markD_02.RData")

# Load cores
no_cores <- parallelly::availableCores() -1

# Load functions
list.files("F:/Analysis/github/project-ettler01/scripts/functions/", pattern="*.*", full.names=TRUE) %>% str_subset( "OBJ", negate = TRUE) %>%  map(~source(.))

# Parameter selection ----
param_sel <- c("initials", "age", "sex", "bmi", "ae_grade_3_4",
               "ae_hyperTAG_grade_3_4","discontinued_due_to_ae", "ae_liver",
               "ae_hemato", "ps_ecog", "stage_early", "first_syst_th",
               "thyroid_disease_before", "monotherapy", "response_achieved",
               "dyslipidemia_before")

param_sel_2 <- c("initials", "age", "sex", "bmi", "ps_ecog", "stage_early",
                 "first_syst_th", "response_achieved", "thyroid_disease_before",
                 "dyslipidemia_before", "monotherapy", "ttnt", "ttnt_achieved",
                 "response_time_to", "response_duration",	"progression",
                 "treatment_duration", "discontinuation_reason", "ae_grade_3_4",
                 "ae_liver", "ae_hemato", "discontinued_due_to_ae",
                 "ae_hyperTAG_grade_3_4", "ae_hyperTAG_any"
                 
)

variab_dep_01 <- c("discontinued_due_to_ae", "ae_hyperTAG_any", 
                    "ae_hyperTAG_grade_3_4", "ae_liver",	"ae_hemato")
variab_indep_01 <- c("age", "sex", "bmi", "ps_ecog", "stage_early", "first_syst_th",
                     "response_achieved", "dyslipidemia_before", 
                     "thyroid_disease_before", "monotherapy"  
)

variab_dep_02a <- c("stage_early", "response_achieved" 
)
variab_dep_02b <- c("treatment_duration", "discontinuation_reason",
                    "ttnt", "ttnt_achieved"
)
variab_dep_02c <- c("response_time_to",                 # only non-zero patients
                    "response_duration", "progression")
variab_indep_02 <- c("age", "sex", "bmi", "ps_ecog", "first_syst_th", "ae_grade_3_4",
                     "monotherapy")
# Object processing  

```


# Questions  
- [ ] In column $response_achieved$ there is value $x$. What does it mean?  
- [ ] The same is in the column $discontinued_due_to_ae$. What does the $x$ mean?  
- [ ] I removed all $lost of FU$.  
- [ ] Should I use $ae_hyperTAG_any$ (column AP) instead of $ae_complete_grade_3_4$ (column AN) in question I.A.2.? In the $ae_hyperTAG_any$ (column AP), there is the selection of any *hyperTAG* in adverse effect.  


# Notes  
* $ae_hyperTAG_any$ is perfectly predicted by $thyroid_disease_before"$. In other words, for that pair of variables, the logistic regression model fits too perfectly — the predicted probabilities are 0 or 1 without uncertainty.


# Exporatory Data Analyses (EDA) {.tabset .tabset-fade .tabset-pills}


## Overview {.unnumbered}
```{r eda_skimr, results="asis"}
d04 |> 
  select(-initials) |>  
  mutate(across(
    .cols = -c(bmi, age), 
    .fns = as.factor
  )) |> 
  my_skim()

```


## Proportional Bar plot {.unnumbered}

```{r eda_barplot}
d04 |> 
  select(-initials, -age, -bmi, -contains("duration"), -ttnt,
         -contains("time_to")) |> 
  select(where(is.numeric)) |> 
  mutate(across(everything(),as.factor)) |> 
  pivot_longer(cols = everything()) |> 
  na.omit() |> 
  ggplot(aes(x = name, fill = as.factor(value))) +
  geom_bar(position = "fill") +  # Proportions
  labs(y = "Proportion", fill = "Outcome") +
  theme_sjplot2()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  paletteer::scale_fill_paletteer_d("waRhol::marilyn_orange_62") +
  NULL
```


# Statistics  
## Multiple correspondence analyses  

```{r stat_mca, out.width="50%"}
res_mca_02 <- d04 |> select(-initials, -age, -bmi) |> na.omit() |> 
  mutate(across(everything(),as.factor)) |> 
  MCA(graph = FALSE)
fviz_mca_var(res_mca_02, 
             select.var = list(contrib = 10),  # only top ten variables
             repel = TRUE, ggtheme = theme_minimal(),
             col.var = "contrib",
             # choice = "mca.cor",
             gradient.cols = c("#bdbdbd", "#ffeda0", "#FC4E07"))

fviz_ellipses(res_mca_02, c("sex", "ae_grade_3_4"),
              geom = "point")
```


## I. section  
### Logistic model  



```{r stat_1_logistic}
res_logist_02 <- d04 |> 
  select(all_of(c(variab_indep_01, variab_dep_01))) |> 
  mutate(sex = if_else(sex == "M", 1, 0)) |> 
  pivot_longer(
    cols = all_of(variab_dep_01),
    names_to = "dep_name",
    values_to = "dep_value"
  ) |> 
  pivot_longer(cols = all_of(variab_indep_01),
               names_to = "indep_name",
               values_to = "indep_value") |> 
  group_by(dep_name,indep_name) |> 
  nest() |> 
  mutate(mod = map(data, ~glm(dep_value ~ indep_value, 
                              data = .x, family = "binomial")),
         tidier = map(mod, ~tidy(.x, conf.int = T, exp = T)))

res_logist_02_tab <- res_logist_02 |> 
  unnest(tidier) |> 
  filter(term == "indep_value") |> 
  select(-data, -mod, -term) |> 
  mutate(across(where(is.numeric), ~round(.x,3)))

res_logist_02_tab |> 
  select(-std.error, - statistic) |> 
  kable(col.names = c("Dependent", "Independent", "Odds Ratio", "p-value",
                      "5% CI", "95% CI")) |> 
  kable_styling(
    full_width = F,
    latex_options = c(
      "hold_position" # stop table floating
    ),
    bootstrap_options = c("striped", "hover", "condensed", "responsive")
  ) %>%
  collapse_rows(columns = 1, valign = "top")
```


## II. section  
### Logistic regression  




```{r stat_2_logistic}
res_logist_03 <- d04 |> 
  select(all_of(c(variab_indep_02, variab_dep_02a))) |> 
  mutate(sex = if_else(sex == "M", 1, 0)) |> 
  pivot_longer(
    cols = all_of(variab_dep_02a),
    names_to = "dep_name",
    values_to = "dep_value"
  ) |> 
  pivot_longer(cols = all_of(variab_indep_02),
               names_to = "indep_name",
               values_to = "indep_value") |> 
  group_by(dep_name,indep_name) |> 
  nest() |> 
  mutate(mod = map(data, ~glm(dep_value ~ indep_value, 
                              data = .x, family = "binomial")),
         tidier = map(mod, ~tidy(.x, conf.int = T, exp = T)))

res_logist_03_tab <- res_logist_03 |> 
  unnest(tidier) |> 
  filter(term == "indep_value") |> 
  select(-data, -mod, -term) |> 
  mutate(across(where(is.numeric), ~round(.x,3)))

res_logist_03_tab |> 
  select(-std.error, - statistic) |> 
  kable(col.names = c("Dependent", "Independent", "Odds Ratio", "p-value",
                      "5% CI", "95% CI")) |> 
  kable_styling(
    full_width = F,
    latex_options = c(
      "hold_position" # stop table floating
    ),
    bootstrap_options = c("striped", "hover", "condensed", "responsive")
  ) %>%
  collapse_rows(columns = 1, valign = "top")
```

### Survival analyses {.tabset .tabset-fade .tabset-pills}  

#### Time to next treatment  {.tabset}  

##### Kaplan-Meier Overall  {.unnumbered}  

```{r stat_2_surv_ttnt_overall}
res_surv_ttnt$surv[[1]]$`KM overall`
```
##### Kaplan-Meier Stratified  {.tabset .unnumbered}  

###### PS-ECOG {unnumbered}  
```{r stat_2_surv_ttnt_stratified_ecog}
res_surv_ttnt$surv[[1]]$`KM stratified`
```

###### First Systemic Therapy {unnumbered}  
```{r stat_2_surv_ttnt_stratified_firstSYSther}
res_surv_ttnt$surv[[2]]$`KM stratified`
```

###### AE Grade 3-4 {unnumbered}  
```{r stat_2_surv_ttnt_stratified_aeGRADE34}
res_surv_ttnt$surv[[3]]$`KM stratified`
```

###### Monotherapy {unnumbered}  
```{r stat_2_surv_ttnt_stratified_monotherapy}
res_surv_ttnt$surv[[4]]$`KM stratified`
```

###### Gender {unnumbered}  
```{r stat_2_surv_ttnt_stratified_sex}
res_surv_ttnt$surv[[5]]$`KM stratified`
```

###### BMI {unnumbered}  
```{r stat_2_surv_ttnt_stratified_bmi}
res_surv_ttnt$surv[[6]]$`KM stratified`
```

###### Age {unnumbered}  
```{r stat_2_surv_ttnt_stratified_age}
res_surv_ttnt$surv[[7]]$`KM stratified`
```





##### Cox Proportional Hazards {.tabset .unnumbered}  

###### PS-ECOG {unnumbered}  
```{r stat_2_surv_cox_table_ecog}
res_surv_ttnt$surv[[1]]$`CoxPH print table`
```

```{r stat_2_surv_cox_plot_ecog}
res_surv_ttnt$surv[[1]]$`CoxPH plot`
```







```{r stat_2_surv_cox_plot}
res_surv_ttnt$surv[[2]]$`CoxPH print table`
```


```{r stat_2_surv_cox_plot}
res_surv_ttnt$surv[[2]]$`CoxPH plot`
```




```{r stat_2_surv_cox_forest_bmi}
res_surv_ttnt$surv[[6]]$`CoxPH forest plot qunatile`
```

```{r stat_2_surv_cox_forest_age}
res_surv_ttnt$surv[[4]]$`CoxPH forest plot qunatile`
```












#### Duration treatment  {.unnumbered}  

```{r stat_2_surv_durTRTeffect}
res_surv_durTRTeffect$surv[[1]]$`KM overall`
```


#### Disease Progression & Response Time  {.unnumbered}  

```{r stat_2_surv_response}
res_surv_response
```



### Linear regression  
#### Time to effect  


```{r}
res_timeTOeffect <- d04 |> 
  select(response_time_to, all_of(variab_indep_02)) |> 
  mutate(sex = if_else(sex == "M",1,0)) |> 
  pivot_longer(cols = -response_time_to,
               names_to = "indep_name",
               values_to = "indep_value") |> 
  group_by(indep_name) |> 
  nest() |> 
  mutate(mod = map(data, ~lm(data = .x, response_time_to ~ indep_value)),
         tidier = map(mod, ~tidy(.x, conf.int = T)),
         qqplot = map2(mod, indep_name, ~ggqqplot(na.omit(resid(.x)) |> as_tibble(),
                                                  x = "value", title = .y)
                       )) 


res_timeTOeffect_tab <- res_timeTOeffect|> 
  unnest(tidier) |> 
  filter(term == "indep_value") |> 
  select(-(data:term), -qqplot)


res_timeTOeffect_tab |> 
  select(-std.error, - statistic) |> 
  kable(col.names = c("Independent", "Estimate", "p-value",
                      "5% CI", "95% CI"),
        digits = 3) |> 
  kable_styling(
    full_width = F,
    latex_options = c(
      "hold_position" # stop table floating
    ),
    bootstrap_options = c("striped", "hover", "condensed", "responsive")
  ) %>%
  collapse_rows(columns = 1, valign = "top")


```


# Conclussion  




# Session info

<details>

<summary>Platform</summary>

```{r session_info_1}
df_session_platform <- devtools::session_info()$platform %>% 
  unlist(.) %>% 
  as.data.frame(.) %>% 
  rownames_to_column(.)

colnames(df_session_platform) <- c("Setting", "Value")

kable(
  df_session_platform, 
  booktabs = T, 
  align = "l",
  caption = "(ref:Reproducibility-SessionInfo-R-environment-title)", # complete caption for main document
  caption.short = " " # "(ref:Reproducibility-SessionInfo-R-environment-caption)" # short caption for LoT
) %>% 
  kable_styling(full_width = F,
                latex_options = c(
                  "hold_position" # stop table floating
                ),
                bootstrap_options = c("striped", "hover", "condensed", "responsive")
  ) 

```

</details>

<details>

<summary>Used packages</summary>

```{r session_info_2}
df_session_packages <- devtools::session_info()$packages %>% 
  as.data.frame(.) %>% 
  filter(attached == TRUE) %>% 
  dplyr::select(loadedversion, date) %>% 
  rownames_to_column

colnames(df_session_packages) <- c("Package", "Loaded version", "Date")

rows <- seq_len(nrow(df_session_packages) %/% 2)
l_session_packages <- list(df_session_packages[rows,1:3],  
                           matrix(numeric(), nrow=0, ncol=1),
                           df_session_packages[-rows, 1:3])

kable(
  l_session_packages, 
  row.names = T,
  booktabs = T, 
  align = "l",
  caption = "(ref:Reproducibility-SessionInfo-R-packages-title)", # complete caption for main document
  caption.short = " " # "(ref:Reproducibility-SessionInfo-R-packages-caption)" # short caption for LoT
) %>% kable_styling(full_width = F,
                    latex_options = c(
                      "hold_position" # stop table floating
                    ),bootstrap_options = c("striped", "hover", "condensed", "responsive") 
) 
```

</details>

<br> <br> <br> <br> <br> <br>

::: {.tocify-extend-page data-unique="tocify-extend-page" style="height: 0;"}
:::

